\input{common/setup}

\begin{document}

\title{Week 1 Practical Session}{Fundamentals of Software Testing}

\begin{enumerate}

    \item Do the ``Who Wants To Be A Software Tester?'' Quiz for this week. If
    you're unclear about any of the answers then ask!

    \item Clone the examples repository (\coderepourl) for the module. Read
    through the {\tt README.md} and ensure that you can get things running on
    Codio or your own machine.

    \item Navigate to the \practicalspackage~package (in the
    \practicalsdirectory~directory of the repository) and open the \weekoneclass
    class. The class contains four methods. Each method has a fault, and you
    will need to write JUnit tests to establish the presence of the fault (by
    producing a failure) and that, following fixing the fault, the method now
    works. Add your tests into the provided skeleton \testweekoneclass JUnit
    test suite class and run them using Maven as described in the code
    repository's \readmefile file.
    
    For each method:

        \begin{enumerate}

            \item Identify the fault.
            
            \item If possible, write a JUnit test case that does {\it not}
            execute the fault.
            
            \item If possible, write a JUnit test case that executes the fault,
            but does not infect the program state.

            \item If possible, write a JUnit test case that causes an infection,
            but {\it not} a failure.

            \item For the given test case, identify the first infected (i.e.,
            erroneous) program state. Be sure to describe the complete state.
            
            \item Fix the fault and verify that the given test now produces the
            expected output.


        \end{enumerate}

\end{enumerate}

\end{document}