\input{common/setup}

\begin{document}

\title{Week 3 Practical Session}{Coverage Part 2}

\begin{enumerate}

    \item Do the ``Who Wants To Be A Software Tester?'' Quiz for this week.
    If you're unclear about any of the answers then ask!

    \item The code segment that returns when a triangle is isosceles could be re-written as:

        \verb$if ((side1 + side2 > side3) &&$\\ 
        \verb$      (side1 == side2 || side2 == side3) &&$\\
        \verb$      (side1 != side2 || slide2 != side3)) {$\\
        \verb$    return Type.ISOCELES;$\\
        \verb$}$

        For the above branch predicate:

        \begin{enumerate}

            \item Enumerate each of the {\it conditions}.
            
            \item Determine the test requirements for coverage with Restricted
            MCDC. Format this as a table, which shows the truth values of the
            conditions and the branch predicate in columns of the table (as done
            in the lecture). Which test requirements are infeasible, if any?

            \item The same question as before, but this time with Correlated MCDC. 
            
            \item Take either (a) your Restricted MCDC test requirements or (b) your
            Correlated MCDC test requirements. Use these to construct a JUnit test
            suite but run against the original \triangleclass class. Note anything
            unusual about the results of your test suite.
            
        \end{enumerate}

    \item Navigate to the \practicalspackage~package (do a {\tt git pull} first), and open the
        \spacedefenderclass class. 
    
            \begin{enumerate}
    
                \item Draw the CCFG for this class.
                
                \item Annotate each node $n$ of the CCFG with $\mathit{defs}(n)$ and 
                $\mathit{uses}(n)$.
    
                \item Write out the set $\mathit{DU}$. In particular, identify any
                elements of $\mathit{DU}$ that are infeasible.
                
                \item Identify which elements of $\mathit{DU}$ are applicable to 
                
                \begin{enumerate}
                    \item All Defs Coverage
                    \item All Uses Coverage
                    \item All Def-Use Paths Coverage
                \end{enumerate}            
                
            \end{enumerate}        
        
\end{enumerate}

\end{document}